Machine Learning Methods for Discourse Segmentation of Communications in E-Mail Based Behavioral Interventions

### 2018-08-07 paper-amia-summit-2018/manuscript.tex
fixed small issues of writing such as gramatical mistakes and spelling errors

The unstructured nature of e-Coaching exchanges poses a unique set of challenges for their qualitative analysis. Usually, email segments are automatically coded with the Minority Youth-Sequential Coding of Process Exchanges (MYSCOPE), \cite{carcone2013provider} a qualitative coding scheme to characterize patient-coach communication during email exchanges. A significant barrier to fully automating the behavior coding process of e-Coaching emails is their segmentation into textual fragments that correspond to distinct communication behaviors. Automating this task is a unique and challenging problem due to the following major reasons:

@@

Relevant previous work in the biomedical domain primarily focused on segmentation of text into sections and headers of electronic health records (EHR)\cite{apostolova2009automatic,denny2009evaluation,tepper2012statistical,cho2002text} or sentence boundary detection of EHR.\cite{griffis2016quantitative,kreuzthaler2015detection,treviso2017sentence} Apostolova et al. applied SVM along with word-vector cosine similarity metric combined with several heuristics to segment clinical reports in EHR into sections, such as demographics, history, procedure, finding and impression.\cite{apostolova2009automatic} After identification of each line in the document, Tepper et al. trained Maximum Entropy models to classify the document into pre-defined sections, such as general patient information, provider information, medical condition, medical history, clinical history, findings, impression, etc.\cite{tepper2012statistical} Denny et al. proposed a SecTag algorithm, which combined natural language processing techniques, terminology-based rules and a Naive Bayes classifier to identify the sections and headers that achieved 99\% recall with 95.6\% precision. \cite{denny2009evaluation} On the other hand, SVM based on prosodic and part of speech features \cite{kreuzthaler2015detection} and recurrent convolutional neural networks using word embeddings \cite{griffis2016quantitative} were utilized for detecting sentence boundaries in EHR. Liu et al. have shown that a linear chain CRF performed better than the hidden markov model and Maxent in sentence boundary detection, evaluated using the official
NIST evaluation tools.\cite{liu2005using} Recently, Treviso et al. proposed a recurrent convolutional neural network for automatic identification of sentence boundary in a neuropsychological text in the Portuguese language.\cite{treviso2017sentence} Their proposed model combined prosodic and part of speech features with pre-trained GloVe word embeddings,\cite{pennington2014glove} which achieved 0.74 F1-score in sentence boundary detection on the Brazilian constitution dataset. Segmentation of e-Coaching emails is different from a traditional shallow discourse analysis of conversations \cite{galley2003discourse} in that the focus is on segmentation, rather than on determining the types of transitions between the utterances or assigning utterances to speakers. Segmentation of e-Coaching emails corresponds to a special type of discourse analysis \cite{webber2012discourse} aimed at better understanding the effective e-Coaching communication strategies and revealing the unique socio-psychological characteristics of a patient.

@@

The experimental dataset for this work was constructed from 49 e-coaching sessions, which include a total of 3,138 segmented and annotated MI behaviors. Each session represents an MI intervention delivered via email. We formulate the segmentation task as a binary classification problem where each word or punctuation mark annotated with a label 1 or 0, to indicate whether it precedes a new segment or not. In total, we obtained 95,777 words and 7,140 punctuation marks, which include 3,138 ``new segment'' and 99,779 ``same segment'' instances. In this study, we experimented with traditional machine learning method, conditional random fields (CRF)\cite{lafferty2001conditional} and deep learning methods, such as multilayer perceptrons (MLP),\cite{rumelhart1986learning} bidirectional recurrent neural networks (bidirectional RNN or simply BRNN)\cite{schuster1997bidirectional} and convolutional recurrent neural networks (CRNN).\cite{treviso2017sentence} For MLP models, samples were created with $2l$ words or punctuations at each step in a word sequence, where a sample contains next $l$ words or punctuations and prior $l$ words or punctuations including current word or punctuation. Each sample is classified into either ``new segment'' or ``same segment'' based on whether the $l^{th}$ word or punctuation in the gold standard precedes a new segment or not. For CRF, BRNN and CRNN models, an email was taken as input sequence, such that POS tags and word embeddings of each word or punctuation mark were used as input and binary labels (1 or 0) corresponding to ``new segment'' and ``same segment'' classification decision were considered as the output of the model at each step. In the gold standard, words or punctuations within the same segment were assigned the label of 0 and the last word or punctuation mark of a segment were assigned the label of 1.

@@

We utilized three types of features in conjunction with CRF, MLP, BRNN and CRNN methods: word embeddings or lexical features, punctuation and POS features. Since POS tags have been shown to be effective semantic abstractions of individual words, we used POS features for our experiment.\cite{liu2005using,treviso2017sentence} To extract our POS features, we tag the e-Coaching emails using the NLTK POS tagger. Punctuation marks, which correspond to one of the symbols \{`.', `,', `!', `?', `:', `;'\} between a pair of words, are also employed as a feature since punctuation marks designate the boundary of a sentence, clause and phrase and often also correspond to a segment boundary.\cite{cho2002text} For natural language processing (NLP) tasks, inputs are received as a text where words can be considered as the basic processing unit. Therefore, it is important to represent a word in such a way that it carries all relevant information. Word embeddings are one of this representation where each word represented as a real-valued vector in a high dimensional vector space. The distributed representation of word allows capturing semantic, syntactic and morphological information from large unannotated corpora.\cite{pennington2014glove, mikolov2013distributed} In our experiment, we utilized embeddings estimated on Google News corpus consisting of 1.6 billion words referred to pre-trained word2vec word embeddings.\footnote{https://code.google.com/p/word2vec/} When words or punctuation marks are not found in the pre-trained word vectors, we utilized word embeddings trained with our e-Coaching email corpus referred to trained word embeddings. CRF utilized lexical features which represent each word in an e-Coaching email.

@@

\textbf{Conditional Random Fields}: CRF has been widely used in various NLP tasks including part-of-speech tagging and segmentation tasks.\cite{lafferty2001conditional, hirohata2008identifying} Unlike a maximum entropy Markov model which uses per-state exponential models for the conditional probabilities of next states given the current state, a CRF is modeled with an exponential distribution provides the conditional distribution of the entire output sequence given the observation sequence. A traditional linear-chain CRF model is defined as a conditional probability distribution $p(y|x)$ for output and input sequences, $y$ and $x$, by the Eq. \ref{Eq:crf}

@@

\textbf{Multilayer Perceptrons}: MLP is a feed-forward artificial neural network which maps set of input onto a set of appropriate output.\cite{rumelhart1986learning} Figure~\ref{fig:mlp} shows a multilayer perceptron with a single hidden layer. In MLP networks, there is no cycles or loops and information moves forward only, from the input nodes through the hidden nodes and to the output nodes. This study utilized MLP with a nonlinear activation function (relu) and one hidden layer consisting of 128 hidden units. In order to prevent over-fitting, we implement dropout on last layer, which randomly hide some neurons during the training phase. \cite{srivastava2014dropout} Dropout is also applied to the fully connected layers in BRNN and CRNN models.

@@

\textbf{Convolutional Recurrent Neural Networks}: CRNN is a deep neural networks architecture,\cite{treviso2017sentence} shown in Figure~\ref{fig:crnn}, consisting of 5 layers: 1) input layer 2) embedding layer 3) convolution layer with max pooling 4) BRNN layer 5) fully connected layer with dropout and sigmoid output. E-coaching email exchanges are represented as a sequence of $m$ words and punctuations, which are fed into the input layer to produce a $m \times n_e$ matrix after fetching the pre-trained word vectors. When POS tags are used in combination with word embeddings, 10-dimensional POS vectors are concatenated with 300-dimensional word vectors to obtain new vectors $n_e = n_w + n_p$ of size 310. The primary purpose of convolution is to extract new features depending on neighbor words. We used 1D convolution which simply moves one step at a time on vertically. In this layer, one filter is responsible for extraction of one feature. After applying $n_f$ different filters with zero-padding on both sides of the text, $n_f$ useful features are produced in the convolution layer for each word. A max pooling operation was then performed over time to find the most significant features in a text region. The bidirectional recurrent layer receives new features extracted in the convolution layer. RNN is usually used to capture the previous history. Moreover, bidirectional RNN is capable of capturing both previous and past histories because bidirectional RNN utilized the features by looking at forward and backward states. The purpose of the fully connected layer is to use the bidirectional RNN layer output for classifying each word or punctuation into ``new segment'' and ``same segment'' classes. Since a fully connected layer has a greater number of parameters, they are more likely excessively co-adapt these parameters causing over-fitting. To prevent this over-fitting problem, dropout is used to randomly drop 50\% of the connections on the fully connected layer. After that, a sigmoid operation is performed on the output layer to provide the probability of whether the word or punctuation precedes a new segment or not. We experimentally determined the optimal parameters and found that the best performance is achieved with 5 folds cross-validation when filter length in convolution layer is 7, number of filters is 100, max-pool size is 3, convolutional layer activation is $relu$, RNN layer activation is $tanh$ and number of recurrent units in RNN is 200. Adam\cite{kingma2014adam} is used for optimization and the early stopping strategy is applied with 50 epochs when the batch size is 32 and the learning rate is 0.001. To utilize our source codes by other research studies, source codes of all experiments will be publicly available\footnote{https://github.com/teanalab/xxxxx} upon acceptance of this paper for publication.

@@

\section*{Results}
Our experimental results have four important directions. First, the optimal dimension of word embeddings and sample size of MLP models are reported. Second, results are reported with respect to ``new segment'' class as well as weighted average over ``new segment'' and ``same segment'' classes in Table~\ref{tab:result_base}. Third, classification performance of different machine learning methods are outlined in Table~\ref{tab:result_weighted_avg} when word embeddings or lexical features are used in combination with punctuation and POS features. Fourth, the impact of the individual, as well as their combination on the segmentation task, are reported.

\begin{figure}[!htb]
    \centering
    \includegraphics[width=1.0\textwidth]{figures/mlp-and-vector.eps}
    \caption{\textbf{Performance of CRNN model on e-Coaching email segmentation by varying the dimension of pre-trained and trained word embeddings with GloVe and word2vec models (left). Performance of MLP model on e-Coaching email segmentation by varying the length of sample size (right).}}
    \label{fig:embedding-dimension-mlp}
\end{figure}

Figure~\ref{fig:embedding-dimension-mlp} (left) illustrates the performance of CRNN model on e-Coaching email segmentation by varying the dimension of pre-trained and trained word embeddings with GloVe\footnote{https://nlp.stanford.edu/projects/glove/} and word2vec models. It was observed that best performance achieved with pre-trained 300-dimensional word2vec word vectors when three types of features are used together. Therefore, we report our results with word2vec 300-dimensional word vectors for all deep learning models used in this study. For MLP models, the first layer input was prepared by the summation of first $l$ word vectors and last $l$ word vectors when a sample contains $2l$ words or punctuations. Figure~\ref{fig:embedding-dimension-mlp} (right) demonstrates the performance of MLP model on e-Coaching email segmentation by varying the value of $l$. It was observed that the best performance of MLP classifier is achieved when $l = 2$. Therefore, results of all MLP models are reported with $l = 2$.

@@

Table~\ref{tab:result_weighted_avg} summarizes the results of all models for segmentation of e-Coaching emails when word embeddings or lexical features are used in combination with punctuation and POS features. Similar to results in Tables~\ref{tab:result_base}, CRNN demonstrates the highest performance among all methods achieving 0.864 recall with 0.825 F1-Measure for ``new segment'' class and 0.990 precision with 0.989 recall and F1-Measure for overall. BRNN and CRF show the lowest and second highest performance, respectively, for email segmentation among all methods. We observed that classification performance significantly improved for ``new segment'' class when word embeddings or lexical features are used in combination with punctuation and POS features. Precision increases by 3.96\%, -2.27\%, 12.71\% and 1.81\%; recall increases by 11.72\%, 19.73\%, 20.59\% and 8.41\%; and F1-measure increases by 8.05\%, 9.51\%, 16.22\% and 5.1\% for CRF, MLP, BRNN and CRNN methods, respectively, in new segment detection when all features are utilized together. Similarly, precision increases by 0.51\%, 0.41\%, 0.82\% and 0.41\%; recall increases by 0.41\%, 0.41\%, 0.72\% and 0.3\%; and F1-measure increases by 0.41\%, 0.41\%, 0.82\% and 0.3\% for CRF, MLP, BRNN and CRNN methods, respectively, in weighted average over ``new segment'' and ``same segment'' classes when word embeddings or lexical features are used in combination with punctuation and POS features.\\

\begin{table}[ht]
\centering
\caption{\textbf{AUC values of all classifiers demonstrating the impact of word embeddings, punctuation and POS features on email segmentation. Highest AUC value for each feature set across all models is highlighted in boldface.}}
\label{tab:result_roc}
 \begin{tabular}{|l|l|l|l|l|}
  \hline
\textbf{Features} & \textbf{CRF} & \textbf{MLP}  & \textbf{BRNN} & \textbf{CRNN} \\ \hline      
 word embeddings only & \textbf{0.966} & 0.951 & 0.920 & 0.965 \\ \hline
 word embeddings + POS & \textbf{0.973} & 0.953 & 0.920 & 0.962 \\ \hline
 word embeddings + punctuation & \textbf{0.993} & 0.989 & 0.982 & 0.988 \\ \hline
 all features & \textbf{0.994} & 0.992 & 0.981 & 0.986 \\ \hline
  \end{tabular}
\end{table}

@@

Table~\ref{tab:result_roc} shows the impact of individual features as well as their combination on the segmentation task. Although CRF provides the second highest classification results, it shows best AUC values, achieving 0.966 AUC when only lexical features are used and 0.994 AUC when a combination of lexical, punctuation and POS features are used. Influence of the punctuation and POS features is also consistent in AUC values, which increase by 2.9\%, 4.31\%, 6.63\% and 2.18\% for CRF, MLP, BRNN and CRNN methods, respectively. Individually, POS and punctuation features improve the performance of all classifiers except CRNN when word embeddings are used with POS features. CRF and MLP achieve their highest AUCs when all features are used together. On the other hand, BRNN and CRNN demonstrate the highest AUCs when word embeddings are combined with punctuation features.

@@

Punctuation mark and POS features made a significant improvement in the performance of machine learning and all deep learning methods. Specially, punctuation features have the higher individual impact on model performance compared to POS features. In all cases, every method performs better, when word embeddings or lexical features are used in combination with punctuation and POS features. This results also indicate that segmentation performance might be improved by adding additional relevant features.

The convolutional layer made a significant difference between the performance of CRNN and BRNN on email segmentation. CRNN demonstrates 22.5\% and 10.7\% higher F1-Measure in ``new segment'' detection and 1\% and 0.5\% higher F1-Measure in overall compared to BRNN when word embeddings and all features are used, respectively. In CRNN, convolution layer performed a series of convolutions and pooling operations during which a number of high-level important features are extracted from the word embeddings of e-Coaching emails. These high-level features are then utilized by the bidirectional RNN of CRNN model, which provide an improved performance on email segmentation. On the other hand, traditional BRNN model received words as input features and their word embeddings are directly utilized in the input layer.

@@

Although punctuation mark plays an important role in segmentation boundary detection, a few errors were encountered by the presence of punctuation marks in boundary identification. For example, a text segment from an e-Coaching email \textit{``A typical day in regards to fruit and vegetable has me eating about a serving at breakfast (our caf has cut up fruit) and then maybe a piece of fruit later in the day or as a snack. Vegetable tends to be a side serving at lunch and dinner and I get celery or carrot cuts with dressing for a snack a lot of times. I could probably add some sort of vegetable into my breakfast (like spinach in an omelet) and snack on another piece of fruit when I am hungry rather than the junk food I tend to eat.''} incorrectly segmented after the first sentence where a punctuation mark was encountered. Similarly, additional information is a common cause for misclassification of an email segment into multiple segments. For instance, the first sentence of the above email segment represents a positive commitment to adolescent's behavior change then next two sentences provide an additional information to support their commitment.

>>>>

The unstructured nature of e-Coaching exchanges poses a unique set of challenges for their qualitative analysis. Usually, email segments are automatically coded with the Minority Youth-Sequential Coding of Process Exchanges (MYSCOPE), \cite{carcone2013provider} a qualitative coding scheme to characterize patient-provider communication during MI session. A significant barrier to fully automating the behavior coding process of e-Coaching emails is their segmentation into textual fragments that correspond to distinct communication behaviors. Automating this task is a unique and challenging problem due to the following major reasons:

@@

Relevant previous work in the biomedical domain primarily focused on segmentation of text into sections and headers of electronic health records (EHR)\cite{apostolova2009automatic,denny2009evaluation,tepper2012statistical,cho2002text} or sentence boundary detection of EHR.\cite{griffis2016quantitative,kreuzthaler2015detection,treviso2017sentence} Apostolova et al. applied SVM along with word-vector cosine similarity metric combined with several heuristics to segment clinical reports in EHR into sections, such as demographics, history, procedure, finding and impression.\cite{apostolova2009automatic} After identification of each line in the document, Tepper et al. trained maximum entropy models to classify the document into pre-defined sections, such as general patient information, provider information, medical condition, medical history, clinical history, findings, impression, etc.\cite{tepper2012statistical} Denny et al. proposed a SecTag algorithm, which combined natural language processing techniques, terminology-based rules and a Naive Bayes classifier to identify the sections and headers that achieved 99\% recall with 95.6\% precision. \cite{denny2009evaluation} On the other hand, SVM based on prosodic and part of speech features \cite{kreuzthaler2015detection} and recurrent convolutional neural networks using word embeddings \cite{griffis2016quantitative} were utilized for detecting sentence boundaries in EHR. Liu et al. have shown that a linear chain CRF performed better than the hidden markov and maximum entropy model in sentence boundary detection, which was evaluated using the official NIST evaluation tools.\cite{liu2005using} Recently, Treviso et al. proposed a recurrent convolutional neural network for automatic identification of sentence boundary in a neuropsychological text in the Portuguese language.\cite{treviso2017sentence} Their proposed model combined prosodic and part of speech features with pre-trained GloVe word embeddings,\cite{pennington2014glove} which achieved 0.74 F1-score in sentence boundary detection on the Brazilian constitution dataset. Segmentation of e-Coaching emails is different from a traditional shallow discourse analysis of conversations \cite{galley2003discourse} in that the focus is on segmentation, rather than on determining the types of transitions between the utterances or assigning utterances to speakers. Segmentation of e-Coaching emails corresponds to a special type of discourse analysis \cite{webber2012discourse} aimed at better understanding the effective e-Coaching communication strategies and revealing the unique socio-psychological characteristics of a patient.

@@

The experimental dataset for this work was constructed from 49 e-Coaching sessions, which include a total of 3,138 segmented and annotated MI behaviors. Each session represents an MI intervention delivered via email. We formulate the segmentation task as a binary classification problem where each word or punctuation mark annotated with a label 1 or 0, to indicate whether it precedes a new segment or not. In total, we obtained 95,777 words and 7,140 punctuation marks, which include 3,138 ``new segment'' and 99,779 ``same segment'' instances. In this study, we experimented with traditional machine learning method, conditional random fields (CRF)\cite{lafferty2001conditional} and deep learning methods, such as multilayer perceptrons (MLP),\cite{rumelhart1986learning} bidirectional recurrent neural networks (bidirectional RNN or simply BRNN)\cite{schuster1997bidirectional} and convolutional recurrent neural networks (CRNN).\cite{treviso2017sentence} For MLP models, instances were created with $2l$ words or punctuations at each step in a word sequence, where an instance contains next $l$ words or punctuations and prior $l$ words or punctuations including current word or punctuation. Each instance is classified into either ``new segment'' or ``same segment'' based on whether the $l^{th}$ word or punctuation in the gold standard precedes a new segment or not. For CRF, BRNN and CRNN models, an email was taken as input sequence, such that POS tags and word embeddings of each word or punctuation mark were used as input and binary labels (1 or 0) corresponding to ``new segment'' and ``same segment'' classification decision were considered as the output of the model at each step. In the gold standard, words or punctuations within the same segment were assigned the label of 0 and the last word or punctuation mark of a segment were assigned the label of 1.

@@

We utilized three types of features in conjunction with CRF, MLP, BRNN and CRNN methods: word embeddings or lexical features, punctuation and POS features. Since POS tags have been shown to be effective semantic abstractions of individual words, we used POS features for our experiment.\cite{liu2005using,treviso2017sentence} To extract our POS features, we tag the e-Coaching emails using the NLTK POS tagger. Punctuation marks, which correspond to one of the symbols \{`.', `,', `!', `?', `:', `;'\} between a pair of words, are also employed as a feature since punctuation marks designate the boundary of a sentence, clause and phrase and often also correspond to a segment boundary.\cite{cho2002text} For natural language processing (NLP) tasks, inputs are received as a text where words can be considered as the basic processing unit. Therefore, it is important to represent a word in such a way that it carries all relevant information. Word embeddings are one of this representation where each word represented as a real-valued vector in a high dimensional vector space. The distributed representation of word allows capturing semantic, syntactic and morphological information from large unannotated corpora.\cite{pennington2014glove, mikolov2013distributed} This study utilized embeddings estimated on Google News corpus consisting of 1.6 billion words referred to pre-trained word2vec word embeddings.\footnote{\label{fn:word2vec}https://code.google.com/p/word2vec/} When words or punctuation marks are not found in the pre-trained word vectors, word embeddings trained with e-Coaching email corpus referred to trained word embeddings were utilized. CRF utilized lexical features which represent each word in an e-Coaching email.

@@

\textbf{Conditional Random Fields}: CRF has been widely used in various NLP tasks including part-of-speech tagging and segmentation tasks.\cite{lafferty2001conditional, hirohata2008identifying} Unlike a maximum entropy markov model which uses per-state exponential models for the conditional probabilities of next states given the current state, a CRF is modeled with an exponential distribution provides the conditional distribution of the entire output sequence given the observation sequence. A traditional linear-chain CRF model is defined as a conditional probability distribution $p(y|x)$ for output and input sequences, $y$ and $x$, by the Eq. \ref{Eq:crf}

@@

\textbf{Multilayer Perceptrons}: MLP is a feed-forward artificial neural network which maps set of input onto a set of appropriate output.\cite{rumelhart1986learning} Figure~\ref{fig:mlp} shows a multilayer perceptron with a single hidden layer. In MLP networks, there is no cycle or loop and information moves forward only, from the input nodes through the hidden nodes and to the output nodes. This study utilized MLP with a nonlinear activation function ($relu$) and one hidden layer consisting of 128 hidden units. In order to prevent over-fitting, we implement dropout on the last layer, which randomly hides some neurons during the training phase. \cite{srivastava2014dropout} Dropout is also applied to the fully connected layers in BRNN and CRNN models.

@@

\textbf{Convolutional Recurrent Neural Networks}: CRNN is a deep neural networks architecture,\cite{treviso2017sentence} shown in Figure~\ref{fig:crnn}, consisting of 5 layers: 1) input layer 2) embedding layer 3) convolution layer with max pooling 4) BRNN layer 5) fully connected layer with dropout and sigmoid output. E-coaching email exchanges are represented as a sequence of $m$ words and punctuations, which are fed into the input layer. The embedding layer provides a distributed representation of words, gives us $m \times n_e$ matrix after fetching the pre-trained word vectors. When POS tags are used in combination with word embeddings, 10-dimensional POS vectors are concatenated with 300-dimensional word vectors to obtain new vectors $n_e = n_w + n_p$ of size 310. The convolution layer used this word embeddings to extract new features depending on neighbor words. This study used 1D convolution which simply moves one step at a time on vertically. In this layer, one filter is responsible for the extraction of one feature. After applying $n_f$ different filters with zero-padding on both sides of the text, $n_f$ useful features are produced in the convolution layer for each word. A max pooling operation was then performed over time to find the most significant features in a text region. The bidirectional recurrent layer receives new features extracted in the convolution layer. RNN is usually used to capture the previous history. Moreover, bidirectional RNN is capable of capturing both previous and past histories because bidirectional RNN utilized the features by looking at forward and backward states. The fully connected layer used the bidirectional RNN output for classifying each word or punctuation into ``new segment'' and ``same segment'' classes. Since a fully connected layer has a greater number of parameters, they are more likely excessively co-adapt these parameters causing over-fitting. To prevent this over-fitting problem, dropout is used to randomly drop 50\% of the connections on the fully connected layer. After that, a sigmoid operation is performed on the output layer to provide the probability of whether the word or punctuation precedes a new segment or not. We experimentally determined the optimal parameters and found that the best performance is achieved with 5 folds cross-validation when filter length in convolution layer is 7, number of filters is 100, max-pool size is 3, convolutional layer activation is $relu$, RNN layer activation is $tanh$ and number of recurrent units in RNN is 200. Adam\cite{kingma2014adam} is used for optimization and the early stopping strategy is applied with 50 epochs when the batch size is 32 and the learning rate is 0.001. To utilize our source codes by other research studies, source codes of all experiments will be publicly available\footnote{https://github.com/teanalab/xxxxx} upon acceptance of this paper for publication.

@@

\section*{Results}
Our experimental results have four important directions. First, the optimal word embedding dimension and instance size of MLP models are illustrated in Figure~\ref{fig:embedding-dimension-mlp}. Second, results are reported with respect to ``new segment'' class as well as weighted average over ``new segment'' and ``same segment'' classes in Table~\ref{tab:result_base}. Third, classification performance of different machine learning methods are outlined in Table~\ref{tab:result_weighted_avg} when word embeddings or lexical features are used in combination with punctuation and POS features. Fourth, the impact of individual features, as well as their combination on the task of email segmentation, are demonstrated in Table~\ref{tab:result_roc}.

\begin{figure}[!htb]
    \centering
    \includegraphics[width=1.0\textwidth]{figures/mlp-and-vector.eps}
    \caption{\textbf{Performance of CRNN model on e-Coaching email segmentation by varying the dimension of pre-trained and trained word embeddings with GloVe and word2vec models (left). Performance of MLP model on e-Coaching email segmentation by varying the length of instance (right).}}
    \label{fig:embedding-dimension-mlp}
\end{figure}

Figure~\ref{fig:embedding-dimension-mlp} (left) shows the performance of CRNN model on e-Coaching email segmentation by varying the dimension of pre-trained and trained word embeddings with GloVe\footnote{https://nlp.stanford.edu/projects/glove/} and word2vec models. It was observed that the best performance of email segmentation was achieved with pre-trained 300-dimensional word2vec word vectors when three types of features are used together. Therefore, results of all deep learning models are reported with pre-trained word2vec 300-dimensional word vectors. In MLP, inputs were prepared by computing the summation of first $l$ word vectors and last $l$ word vectors where each instance contains $2l$ words or punctuations. Figure~\ref{fig:embedding-dimension-mlp} (right) demonstrates the performance of MLP models on e-Coaching email segmentation by varying the value of $l$. It was observed that the best performance of MLP classifiers is achieved when $l = 2$. Therefore, results of all MLP models are reported with $l = 2$.

@@

Table~\ref{tab:result_weighted_avg} summarizes the results of all models for segmentation of e-Coaching emails when word embeddings or lexical features are used in combination with punctuation and POS features. Similar to results in Tables~\ref{tab:result_base}, CRNN demonstrates the highest performance among all methods achieving 0.864 recall with 0.825 F1-Measure for ``new segment'' class and 0.990 precision with 0.989 recall and 0.989 F1-Measure for overall. BRNN and CRF show the lowest and second highest performance, respectively, for email segmentation among all methods. We observed that classification performance significantly improved for ``new segment'' class when word embeddings or lexical features are used in combination with punctuation and POS features. Precision increases by 3.96\%, -2.27\%, 12.71\% and 1.81\%; recall increases by 11.72\%, 19.73\%, 20.59\% and 8.41\%; and F1-measure increases by 8.05\%, 9.51\%, 16.22\% and 5.1\% for CRF, MLP, BRNN and CRNN methods, respectively, in new segment detection when all features are utilized together. Similarly, precision increases by 0.51\%, 0.41\%, 0.82\% and 0.41\%; recall increases by 0.41\%, 0.41\%, 0.72\% and 0.3\%; and F1-measure increases by 0.41\%, 0.41\%, 0.82\% and 0.3\% for CRF, MLP, BRNN and CRNN methods, respectively, in weighted average over ``new segment'' and ``same segment'' classes when word embeddings or lexical features are used in combination with punctuation and POS features.\\

\begin{table}[ht]
\centering
\caption{\textbf{AUC values of all classifiers demonstrating the impact of word embeddings, punctuation and POS features on e-Coaching email segmentation. Highest AUC value for each feature set across all models is highlighted in boldface.}}
\label{tab:result_roc}
 \begin{tabular}{|l|l|l|l|l|}
  \hline
\textbf{Features} & \textbf{CRF} & \textbf{MLP}  & \textbf{BRNN} & \textbf{CRNN} \\ \hline      
 word embeddings only & \textbf{0.966} & 0.951 & 0.920 & 0.965 \\ \hline
 word embeddings + POS & \textbf{0.973} & 0.953 & 0.920 & 0.962 \\ \hline
 word embeddings + punctuation & \textbf{0.993} & 0.989 & 0.982 & 0.988 \\ \hline
 all features & \textbf{0.994} & 0.992 & 0.981 & 0.986 \\ \hline
  \end{tabular}
\end{table}

@@

Table~\ref{tab:result_roc} shows the impact of individual features as well as their combination on the segmentation task. Although CRF provides the second highest classification results, it shows best AUC values, achieving AUC 0.966 when only lexical features are used and AUC 0.994 when a combination of lexical, punctuation and POS features are used. Influence of the punctuation and POS features is also consistent in AUC values, which increase by 2.9\%, 4.31\%, 6.63\% and 2.18\% for CRF, MLP, BRNN and CRNN methods, respectively. Individually, POS and punctuation features improve the performance of all classifiers except CRNN when word embeddings are used with POS features. CRF and MLP achieve their highest AUCs when all features are used together. On the other hand, BRNN and CRNN demonstrate the highest AUCs when word embeddings are combined with punctuation features.

@@

Punctuation mark and POS features made a significant improvement in the performance of traditional machine learning and deep learning methods. Specially, punctuation features have the higher individual impact on model performance compared to POS features. In all cases, every method performs better, when word embeddings or lexical features are used in combination with punctuation and POS features. This results also indicate that segmentation performance might be improved by adding additional relevant features.

The convolutional layer made a significant difference between the performance of CRNN and BRNN on email segmentation. CRNN demonstrates 22.5\% and 10.7\% higher F1-Measure in ``new segment'' detection and 1\% and 0.5\% higher F1-Measure in overall compared to BRNN when word embeddings and all features are used, respectively. In CRNN, convolution layer performed a series of convolutions and pooling operations during which a number of high-level important features are extracted from the word embeddings of e-Coaching emails. These high-level features are then utilized by the bidirectional RNN of CRNN model, which provide an improved performance on email segmentation. On the other hand, traditional BRNN model directly utilized word embeddings at the bidirectional RNN layer.

@@

Although punctuation mark plays an important role in segmentation boundary detection, a few errors were encountered by the presence of punctuation marks in boundary identification. For example, a text segment from an e-Coaching email \textit{``A typical day in regards to fruit and vegetable has me eating about a serving at breakfast (our caf has cut up fruit) and then maybe a piece of fruit later in the day or as a snack. Vegetable tends to be a side serving at lunch and dinner and I get celery or carrot cuts with dressing for a snack a lot of times. I could probably add some sort of vegetable into my breakfast (like spinach in an omelet) and snack on another piece of fruit when I am hungry rather than the junk food I tend to eat.''} incorrectly segmented after the first sentence where a punctuation mark was encountered. Similarly, additional information is a common cause for misclassification of an email segment into multiple segments. For example, the first sentence of the above email segment represents a positive commitment to adolescent's behavior change then next two sentences provide an additional information to support their commitment.
